\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-03-30 redaction/dhp_july2017.tex
changement papier

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs. In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects. This drives us to provide a careful modeling of international transport costs, that allows for both ad-valorem (multiplicative) and per-unit (additive) components, exploiting information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. In contrast to \cite{hummels2007}, trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the additive component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}). This view does not make consensus though. Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques should be compared with technical progress in production to get a clear picture of the transport costs burden. The constant-dollar cost of transport has declined, but so has the constant-dollar cost of production. As well, the quality of transport services has increased. In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries). Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs, suggesting that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip

@@

In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database. This focus is shared with part of the international trade literature, such as \cite{hummels2007} and \cite{Behar_Venables}. Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). There is no clear evidence about this though, in particular when expressing the transport cost burden in terms of the export price. According to \cite{hummels2007}, transport costs in proportion of export prices have thus declined less quickly than expected. His results point as an important explanation, that the reduction in pure transport costs in both maritime and air shipping is partially offset by trade composition effects, thereby attenuating the downward trend in overall (observed) transport costs. We now investigate the robustness of this result. \textbf{stop here}.

@@

Two main comments can be made. First, in both transport modes, the additive component appears of sizeable importance all over the period, as the share of additive costs remains roughly constant over time. Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs.

Second, considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}, with a magnitude of transport costs that appears higher in ocean shipping than for air transport over the period.\smallskip

Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both \textit{i)} the evolution of per product and per partner ``ceteris paribus'' transport costs and \textit{ii)} the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased (explanation \textit{ii)}), independently of any change in transport costs \textit{per se}. As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of ``per se'' or ``ceteris paribus'' international transport costs, i.e. per product- and per partner- transport costs.

\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}. He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f$ the shipping charge per kg shipped. Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, the transport cost measure writes down as:

@@

Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value). As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem. At the other extreme, $\beta = 1$ means that all transport costs are additive. Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs. Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which Hummels (2007) decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}.

In doing so, Hummels (2007) implicitly assumes $\beta$ constant across the triplet origin country origin/product/year. In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions. However, our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:decomp_TC_3d}, cast some doubt on the empirical relevance of this assumption. As reported in Figure \ref{fig:decomp_TC_3d}, if overall transport costs have decreased over time, it is also the case for each additive and multiplicative components. In particular, the share of additive costs appears to be varying over time, in both transport modes. On light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country.

@@

As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending of the transport mode, consistently with our previous findings. This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would rather imply a distribution concentrated on a single point. Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year. These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and transport costs ``per se''.

@@

Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country. Starting from these values, we extract the ``ceteris paribus'' transport cost measure, for each additive and multiplicative cost component (by transport mode) by the mean of a time fixed effect. Precisely, starting from the values of each additive and ad-valorem transport cost measure estimated in Section \ref{sec:results_decomposition}, we extract the changes over time in the ``ceteris paribus'' transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components.
For each cost component (additive and ad-valorem), comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the ``ceteris paribus'' costs (for instance, insurance or handling costs) that have reduced over time. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the ``unfitted'' total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).}

The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and overall resp.), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and overall respectively). We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).

@@

\paragraph{Obtaining the ``pure'' transport costs component series} For each additive and multiplicative transport costs component, the empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process. First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects. For the estimated ad-valorem component, we thus estimate the following equation:
\begin{eqnarray}
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

@@

%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}
As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), the objective is to decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the ``ceteris paribus'' transport costs time trend (Term (c)). Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions. Both equations are estimated by transport mode.\smallskip

@@

\paragraph{Obtaining the unfitted measures} The objective here is to get the unfitted transport cost component (additive and multiplicative) as as index, with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013). That is, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:

@@

We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the ``pure'' transport cost series (i.e., composition effects excluded). As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974. Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain. \smallskip

\paragraph{Obtaining the unfitted total transport cost index} For the unfitted measure, we build for each transport mode, the ``total'' transport cost based on Equation (\ref{eq:base_estimee}), according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \widehat{t}_t$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ have been estimated (by year) as explained in Section \ref{sec:data_method}. Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\widehat{t}_t$ the additive component, both expressed in percentage of the fas price. We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though given that the fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices. In a first step then, we re-build the fitted measures of each transport cost component in value starting from these indices. For the ad-valorem component, this can be obtained making use of Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:

$$\widehat{\tau}^{cp, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$

with $\widehat{\tau}^{cp, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) ad-valorem cost. A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:

$$\widehat{t}^{cp}_t = \frac{\Gamma^{add}_t}{100}$$

with $\widehat{t}^{cp}_t \equiv \exp(\gamma_t)$ the yearly value of the \textit{ceteris paribus} (fitted) additive cost.

We then deduce the fitted value of the overall cost according to:
$$\widehat{tc}^{cp}_t= \widehat{\tau}^{cp, adv}_t  + \widehat{t}^{cp}_t$$

Last, we transform this (fitted) ``overall'' transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{cp}_t -1 }{\widehat{tc}^{cp}_{1974}-1}$$

@@

%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
\begin{table}[htbp]
  \centering
  \caption{Air: Transport costs estimates, 3-digits (selected years)}
\begin{center}
    \begin{tabular}{l|ccccccc}
\hline\hline
Year & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
\hline
\multicolumn{8}{l}{\textbf{Model (A) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}$, in \%)}     \\
\hline
Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.0 & 4.2 & 3.4  \\
Median & 5.4 & 3.8 & 4.4 & 2.5 & 2.9 & 3.4 & 2.9  \\
%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
\hline
\multicolumn{8}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem TC}}    \\
\hline
\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}$, in \%)}  \\ \hline
Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
Median & 2.7 & 1.6 & 1.6 & 1.2 & 1.4 & 2.2 & 1.7 \\
%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\

>>>>

\begin{abstract}
This paper investigates the underlying sources behind the downward trend of international transport costs. In the same spirit as \cite{hummels2007}, we identify the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, using information contained in the US imports flows database over 1974-2013 distinguished by transport mode, air and vessel. In contrast to \cite{hummels2007} though, we find that trade composition effects do not matter much and when they do, they tend to amplify (rather than reduce) the decrease in pure transport costs. Importantly, this difference of results can be attributed to the new way of modeling the per-unit (additive) component of transport costs we offer. Further, in quantitative terms, we find that the additive component of transport costs is sizeable, representing between 40 and 50\% of overall transport costs. In both aspects, our results point the importance of the additive component in accounting for international transport costs.
\end{abstract}

@@

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation (\citealp{levinson2016box}), infrastructure development and new communication technologies (\citealp{Lafourcade_Thisse}). This view does not make consensus though. Distance as an incompressible cost is still a very important determinant of trade flows (\textbf{citer?}); further, progress in transportation techniques (that reduces the constant-dollar cost of transport) should be compared with technical progress in production (that reduces the constant-dollar cost of production) to get a clear picture of the evolution of the transport costs burden. In accordance with this view, \cite{hummels2007} points out that transport costs in proportion of export prices have declined less quickly than expected.\smallskip

According to \cite{hummels2007}, this lower-than-expected decline of transport costs is partly explained by trade composition effects (i.e., changes over time in the baskets of imported goods and/or the origin countries). Based on US imports over 1974-2004, he finds that total transport costs have decreased in both air and maritime shipment, but less so than ``pure'' transport costs (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) because trade composition effects have partly offset the ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend.\footnote{Throughout the paper, we refer to ``pure'' or ``per se'' transport costs changes as the changes over time in the transport costs for a given product/country partner, ie excluding trade composition effects.} In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the evolution of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs. In this respect, our results suggest that pure transport costs declined even less than argued by \cite{hummels2007}.\medskip

@@

In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database. This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}. Specifically, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, in the same spirit as \cite{hummels2007}.

@@

Two main comments can be made. First, in both transport modes, the additive component appears of sizeable importance all over the period. Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs. However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip

Second, considering the trend in total transport costs (the upper line in Figure \ref{fig:decomp_TC_3d}), both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip

Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased , independently of any change in transport costs \textit{per se}. As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs. Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974. Our own results tend to challenge this view, as we further show.

\paragraph{A word on Hummels' methodology} At this point, let us make a brief summary of Hummels' methodology\nocite{hummels2007}. He starts from the following specification of the ratio of destination to origin prices modelled as $p_{ikt} = \widetilde{p}_{ikt}+\frac{f_{ikt}}{ \widetilde{p}_{ikt}},$ with $f_{ikt}$ the shipping charge per kg shipped (for a given product $k$ imported from country $i$ in year $t$). Further, and as in \cite{hummels_skiba}, \cite{hummels2007} writes the per kg shipping charge as $f_{ikt}=\widetilde{p}_{ikt}^{1-\beta}X_{ikt}$, where $X_{ikt}$ represents other costs shifters (distance, port quality, etc.) Accordingly, his transport cost measure $TC_{ikt}$ writes down as:

@@

Notice that $\beta$ can be interpreted as the elasticity of transport costs to the export price (in absolute value). As such, a value of $\beta = 0$ corresponds to the standard case of ``iceberg'' costs, in which transport costs are purely ad valorem. At the other extreme, $\beta = 1$ means that all transport costs are additive. Accordingly, the estimated value of $\beta$ can be interpreted as the share of additive costs in total transport costs. Equation (\ref{eq:Hummel_benchmarkeq}) is the baseline specification from which \cite{hummels2007} decomposes the changes in transport costs over time between trade composition effects and changes in ``pure'' transport costs, as described with more details in Appendix \ref{app:compare_Hummels}. In doing so, \cite{hummels2007} implicitly assumes $\beta$ constant across the triplet origin country origin/product/year. In light of the economic interpretation of $\beta$, this means that the share of additive costs in total costs does not vary across these dimensions.

\paragraph{Challenging the $\beta$'s constancy assumption} Our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:part_cout_additif}, cast some doubt on the empirical relevance of this assumption. As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes. In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.

@@

As displayed in Figure \ref{fig:histogram_beta}, for both transport modes the distribution of $\beta_{ikt}$ over the triplet (year, product, origin country) is smoothly distributed over the interval $[0,1]$, with a mode of the distribution standing around 0.3-0.4 depending of the transport mode, consistently with our previous findings. This stands in sharp contrast with the assumption made by  \cite{hummels_skiba} or \cite{hummels2007}, which would rather imply a distribution concentrated on a single point. Rather, this suggests that the elasticity of transport costs to the export price, or equivalently the share of additive costs in total costs, is varying across time, product and country partner.\footnote{We have checked that it is also the case when we look at the distribution over the range of origin countries or products, for a given year. These results are not reported for sake of space saving but they are available upon request to the authors.} In the next section, we shed light on the role of this result in the analysis of the underlying sources of the overall transport costs trend patterns, between trade composition effects and pure transport costs.

@@

Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$). Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect. Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).

One advantage of this method is to yield measures of transport costs (fitted and unfitted) that are easily comparable between transport modes and transport cost components. For each cost component (additive and ad-valorem) as well as for the total transport cost, comparing the unfitted measure and the fitted measure (composition effects excluded) allows to characterize if the decrease observed over the period in the unfitted series is due to trade composition effects (for instance, changes in the country partners, in the type/ quality of products traded), or if it is the pure transport costs (for instance, insurance or handling costs) that have reduced over time. We now describe the method to extract these series (with more details provided in Appendix \ref{app:comp-effects}).

@@

\paragraph{Obtaining the ``pure'' transport costs component series} The empirical strategy to get the fitted transport cost measure (i.e., extracting from trade composition effect) can be described as a two-stage process. First, we decompose the estimated measure in the three product/country/time dimensions, using fixed effects. For the estimated ad-valorem component, we estimate the following equation:
\begin{eqnarray}
\ln(\widehat{\tau}_{ikt})&=&\delta +\underbrace{\sum_{i \neq \text{ARG}}\alpha_i.\mathbbm{1}_i}_{(a)} + \underbrace{\sum_{s(k)\neq \text{011}}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)} + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_mult}
\end{eqnarray}

@@

%, based on the additive cost series $\widehat{t}_{ikt}$ previously obtained:
\begin{eqnarray}
\ln(\widehat{t}_{ikt})&=&\ln\left( \delta + \underbrace{\sum_{i \neq \text{ARG}}  \alpha_i.\mathbbm{1}_i}_{(a)}+\underbrace{\sum_{s(k)}\beta_{s(k)}.\mathbbm{1}_{s(k)}}_{(b)}\right) + \underbrace{\sum_{t \neq 1974}\gamma_t.\mathbbm{1}_t}_{(c)}+\epsilon_{ikt} \label{eq:compeffects_add}
\end{eqnarray}
As displayed in Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}), we decompose the estimated transport cost component in three elements: the country dimension (Term (a)), the product dimension (Term (b)) and the pure transport costs time trend (Term (c)). Note that Equations (\ref{eq:compeffects_mult}) and (\ref{eq:compeffects_add}) preserve our specification of the ad-valorem and additive costs of Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}), as we consider that the iceberg cost is the product of the country of origin and the good dimension, while the additive cost is the sum of the two dimensions. Both equations are estimated by transport mode.\smallskip

@@

\paragraph{Obtaining the unfitted measures} The objective is to get the unfitted transport cost component (additive and multiplicative) built as as index with reference value 100 in 1974, starting from the estimated values previously obtained ($\widehat{\tau}_t$, $\widehat{t}_t$, for each year $t$ from 1974 to 2013). To do so, we apply the simple formula to get the following indices, for the ad-valorem and the additive cost components respectively:

@@

We also build two measures of the ``total'' transport costs, that agglomerates our estimates of the two additive and ad-valorem components, for both the unfitted series and the pure transport cost series (i.e., composition effects excluded). As for each additive and ad-valorem component, the objective is to get a measure of total transport cost changes built as an index starting from the value 100 in 1974. Even if obeying to the same logic, we proceed slightly differently for the unfitted and the fitted measures, as we now explain. \smallskip

\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the ``total'' transport cost based on Equation (\ref{eq:base_estimee}) according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}. Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed in percentage of the fas price. We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

\paragraph{Obtaining the fitted total transport cost index} The same logic as above applies to construct the fitted measure of total transport cost (i.e., composition effect excluded), with one notable exception though: Fitted ad-valorem and additive components have not been estimated in value, but extracted and build as indices. In a first step then, we re-build the fitted measures of each transport cost component in value starting from these indices. For the ad-valorem component, this can be obtained making use of Equation (\ref{eq:tcadv_compoeffect}), rewritten to get:

$$\widehat{\tau}^{pure, adv}_t = \frac{\Gamma^{adv}_t \left(\bar{\tau}_{1974}-1\right)}{100} +1$$

with $\widehat{\tau}^{pure, adv}_t \equiv \bar{\tau}_{1974}.\exp(\gamma_t)$ the yearly value of the pure (fitted) ad-valorem cost. A similar reasoning starting from Equation (\ref{eq:tcadd_compoeffect}) gives the fitted value for the additive cost component as:

$$\widehat{t}^{pure}_t = \frac{\Gamma^{add}_t}{100}$$

with $\widehat{t}^{pure}_t \equiv \exp(\gamma_t)$ the yearly value of the pure (fitted) additive cost.

We then deduce the fitted value of the overall cost according to:
$$\widehat{tc}^{pure}_t= \widehat{\tau}^{pure, adv}_t  + \widehat{t}^{pure}_t$$

Last, we transform this (fitted) measure of total transport cost in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc}_t = 100\frac{\widehat{tc}^{pure}_t -1 }{\widehat{tc}^{pure}_{1974}-1}$$

\textbf{stop here}

@@

%\todo{Changer la facon de presenter les resultats, pour harmoniser avec la facon de faire dans le corps du papier. Jerome?}
\begin{table}[htbp]
  \centering
  \caption{Air: Transport costs estimates, 3-digits (selected years)}
\begin{center}
    \begin{tabular}{l|ccccccc}
\hline\hline
Year & 1974  & 1980  & 1990  & 2000  & 2004 & 2010  & 2013   \\
\hline
\multicolumn{8}{l}{\textbf{Model (a) - With only Ad-Valorem TC} ($\widehat{\tau}^{ice}-1$, in \%)}     \\
\hline
Mean  & 6.9& 5.4 &5.0 & 3.6 & 4.0 & 4.2 & 3.4  \\
Median & 5.4 & 3.8 & 4.4 & 2.5 & 2.9 & 3.4 & 2.9  \\
%Standard Error & 0.052 & 0.049 & 0.039 & 0.033 & 0.037 & 0.024 \\
\hline
\multicolumn{8}{l}{\textbf{Model (b) - With Additive \& Ad-Valorem TC}}    \\
\hline
\multicolumn{8}{l}{\textit{Ad-valorem term} ($\widehat{\tau}^{adv}-1$, in \%)}  \\ \hline
Mean & 3.6 & 2.3 & 2.4 &1.7 & 1.9 & 2.6 & 1.7  \\
Median & 2.7 & 1.6 & 1.6 & 1.2 & 1.4 & 2.2 & 1.7 \\
%Standard Error & 0.032 & 0.025 & 0.021 & 0.016 & 0.023 & 0.012  \\
